---
title: "Writing Functions"
author: "Fonti Kar"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A quick explaininer on writing R functions

As researchers we often need to repeat our code operations. This often results in a lot of copy and pasting, making our R scripts/files realllly long. This also creates a problem of having to make changes to all your repeated code if you need to edit one thing. You can see how quickly things can get out of hand....

This is where functions can be useful. You write a function to do your operation!  This not only reduces your copy and paste load, but it also means you only need only change your code in your function **once** if you need to make any changes. 

Here's an example: 

We need to reach each penguin dataset and clean it. I've done it manually for 3 species and already you can see how much copy and pasting that was! 

## Read in af data

```{r}
af <- read_csv("data/penguin_distributions/Aptenodytes forsteri.csv")
```

## Clean af data

```{r}
af |> 
  select(gbifID:species, individualCount,
         countryCode, locality, 
         institutionCode,
         decimalLongitude, decimalLatitude,
         eventDate, 
         coordinateUncertaintyInMeters) |> 
  filter(individualCount > 1,
         institutionCode %in% c("CLO", "iNaturalist")
         )|>
  arrange(eventDate) |> 
  clean_names() |> 
  write_csv("output/af.csv")
```

## Read in ap data

```{r}
ap <- read_csv("data/penguin_distributions/Aptenodytes patagonicus.csv")
```

## Clean ap data

```{r}
cleaned_ap <- ap |> 
  select(gbifID:species, individualCount,
         countryCode, locality, 
         institutionCode,
         decimalLongitude, decimalLatitude,
         eventDate, 
         coordinateUncertaintyInMeters) |> 
  filter(individualCount > 1,
         institutionCode %in% c("CLO", "iNaturalist")
         )|>
  arrange(eventDate) |> 
  clean_names() 

cleaned_ap |> 
  write_csv("output/ap.csv")
```

## Read in ec data

```{r}
ec <- read_csv("data/penguin_distributions/Eudyptes chrysocome.csv")
```

## Clean ap data

```{r}
cleaned_ec <- ec |> 
  select(gbifID:species, individualCount,
         countryCode, locality, 
         institutionCode,
         decimalLongitude, decimalLatitude,
         eventDate, 
         coordinateUncertaintyInMeters) |> 
  filter(individualCount > 1,
         institutionCode %in% c("CLO", "iNaturalist")
         )|>
  arrange(eventDate) |> 
  clean_names() 

cleaned_ec |> 
  write_csv("output/ec.csv")
```

## Function to prep data

So here I've decided that was too much work and I need to write a function. I've called my function `prepare_data_for_plot`. Give your functions a clear, declarative name of what it does. Not `my_function`, something meaningful. 

- To write your function, we use the `function()` function. The inputs you put inside `function()`, are the inputs you will have for your function. Here, my function will take an input `function(data)`. 

- The `{    }` is where you define the body of your function. This is where I've pasted my repeated data cleaning code. 

Notice how I've assigned the output of my cleaning pipe to `cleaned_data` and at the end of my function I've called the `return(cleaned_data)` so that when I call my `prepare_data_for_plot()` it will also print the `cleaned_data` object into the console 

```{r}
prepare_data_for_plot <- function(data){
  
  cleaned_data <- data |> 
  select(gbifID:species, individualCount, # Selecting specific cols
         countryCode, locality, 
         institutionCode,
         decimalLongitude, decimalLatitude,
         eventDate, 
         coordinateUncertaintyInMeters) |> 
  filter(individualCount > 1, # Exclude records with less 1 individual count
         institutionCode %in% c("CLO", "iNaturalist")
         )|>
  arrange(eventDate) |>  # Sort by date
  clean_names()  # clean colname names

  return(cleaned_data)
  
}

cleaned_af <- prepare_data_for_plot(af)
prepare_data_for_plot(ap)
prepare_data_for_plot(ec)

```
