---
title: "Class Notes on Reproducible Research"
output: html_document
date: "2024-07-24"
editor_options: 
  chunk_output_type: console
---


# What is ideal reproducible research?

- Transparent methods
- Accessible data and code
- Can be replicated
- Interoperable across operating systems and contexts

Historically, science was secretive but there has been a recent push to make science more open.

Many aspects of science can be made transparent, open, replicable. This course is about computation reproducibility.

Goal as an open scientist: Be able to reproduce the results at a later time, better if collaborator can do so, someone from research community.

There are processes to handle sensitive data (keep the science part open but the sensitive parts, private)

# Activity: Identify challenges/problems/barriers for reproducible analyses and collaboration

(see detailed notes in attached file)

- Software environment
- Workflow
- Coding conventions
- Versions
- Documentation
- Data management
- Big data

# Organisation

Projects

- a project is a sensible unit of organisation for your work. Usually this is a paper, PhD chapter, or report
- everything you need for your project is in this folder
- each project should be an separate github repository
- Add an Rstudio project file to the project, and use this to open Rstudio in the right folder. See https://r4ds.had.co.nz/workflow-projects.html#rstudio-projects

Have predictable folder structure within your projects

- ideally align with community practices. 
- a common folder structure is something like this
```
my_porject
├── my_porject.Rproj      # Rstudio project file
├── README.md             # a text file with high level summary of the project
├── analysis.Rmd          # main file running your analysis
├── data                  # raw data files & metadata
├── downloads             # data files that are downloaded
├── docs                  # any documents you're stroing
├── ignore                # a folder to dump stuff in that is ignored by git
├── R                     # folder with custom R functions
├── outputs               # folder for figures or files genearted by your code
└── scripts               # R scripts 
```

Within the project, use relative paths in your code

- Avoid absolute pathnames, as these break when the folder is moved within your computer or to your collaborators computer

```{r}
# Yes!!!! do this
penguin_sizes <- read.csv("data/penguin_sizes.csv")

# No!!!! Avoid  this
penguin_sizes <- read.csv("/Users/dfalster/GitHub/projects/2024_R_intermediate_UNSW/data/penguin_sizes.csv")
```

Include an obvious landing point in your code. 

- i usually add a file called `Analysis.Rmd` in the project root, as an obvious point to go
- extra details can go in the folder called `README.md`

# Data curation and storage

```{r}

```



# R / Rmd / Qmd

### .R 
Traditional we have written our R code in `.R` files. These are great, especially for simple analyses or for storing functions (more on these tomorrow!). In `.R` files, we would write our comments using  the hash tag e.g. `# This is a comment`. Lines with `#` at the start will be treated as comments and R won't read those lines as code. Comments are handy when we want to give ourselves/reader a little reminder what is happening at each line but if you want to elaborate or talk about the maths/stats, comments are a bit clunky!

### .Rmd
In situations where you want to elaborate alongside your code, `.Rmd` (RMarkdown) is a great option! `.Rmd` allows you to embed R code/figures/hyperlinks/tables with regular plain text. Don't worry you can still use comments. Within an `.Rmd`, your R code will live in a code chunk (grey boxes). The great thing about `.Rmd` is that you can knit/render the file into a beautiful output (`.html`/`.docx`/`.pdf`). Check out this [intro](https://rmarkdown.rstudio.com/articles_intro.html).

[RMarkdown](https://rmarkdown.rstudio.com/index.html) allows you to create really beautiful reports, you can even make slides as well. A lot of the fine tuning comes to the header of the `.Rmd`. For example, you can add an author field.

```
---
title: "Penguin sizes"
output:
  html_document: default
date: "2024-07-23"
authors: Daniel Falster, Fonti Kar, Will Cornwell
--
```

Some code will print extra messages e.g `library(tidyverse)`. You can control the behaviour of the output your R chunks `r, message=FALSE}`. Setting `message=FALSE` will turn off the messages that come from your R code. Learn more about the chunk options [here](https://rmarkdown.rstudio.com/lesson-3.html) 

### .Qmd

Quarto documents (`.Qmd`), is a new age document file type. It is built from `.Rmd`. If you know how to use `.Rmd`, you will know how to use a `.Qmd`. The biggest different is that `.Qmd` allows for cross-language collaborations.  Python and R users can contribute to the same document. 

There is an interactive editor but this is a little buggy

You will need to [download Quarto](https://quarto.org/docs/get-started/) in order to use Quarto docs. 

Quarto is really powerful, it allows you to build books, websites, blogs as well as beautiful rendered reports. 


# Reproducible coding

The biggest keys to making your code reproducible are

1. Start with a script that has runs end to end
2. Vigoroursly maintain this as the project develops!

Sounds easy!

some advice on maintaining the reproducibility include

- removing old or broken code 
- disabling Rstudio's bad default habit of saving and reloading objects from the working environment (see here on how to do it)
  - see https://r4ds.had.co.nz/workflow-projects.html#what-is-real
  - this means you have to rerun your code each time you reopen the project. This quickly uncovers any issues that arise
  - for most projects, the cost of rerunning the analysis is low
- as soon as you encounter an issue, fix it! The work to keep your script in full working order is low. The work to upgrade a script that has layers of issues is high.

If you have slow running steps (.e.g more than 1min), you can use an if statement, which only runs the slow code if the file is missing.

```{r}

path_model_outputs <- "outputs/model_runs.rds"

if(!file.exists(path_model_outputs)) {
  
  # beware, takes 4 hrs
  fits <- run_my_models(data)
  
  savesRDS(fits, path_model_outputs)
}

outputs <- readRDS(path_model_outputs)
```

You can also use the above approach if you have manual steps.


# Saving outputs

```{r}
library(tidyverse)
library(ggplot2)
library(maps)
library(ggthemes)

af <- read_csv("data/penguin_distributions/Aptenodytes forsteri.csv")

# Get world map data
world_map <- map_data("world")

af$countdatapresent <- case_when (af$individualCount > 0 ~ "Count data present", .default = "No count data")

#mostly from chagpt:
# Create a ggplot object
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "white") + # Fill the regions and outline with white
  scale_fill_viridis_d() + # Use a viridis color scale for regions
  theme_minimal() + # Apply a minimal theme
  theme(
    panel.background = element_rect(fill = "lightblue"),
    # Light blue background
    panel.grid.major = element_blank(),
    # Remove major grid lines
    panel.grid.minor = element_blank(),
    # Remove minor grid lines
    axis.title = element_blank(),
    # Remove axis titles
    axis.text = element_blank(),
    # Remove axis text
    axis.ticks = element_blank() # Remove the legend
  ) +
  coord_fixed(1.3) +
  geom_point(data = af,
             aes(
               x = decimalLongitude,
               y = decimalLatitude,
               group = NA,
               col = countdatapresent
             ))# Ensure the aspect ratio is correct

ggsave("output/emperorpenguindist.pdf",
       width = 5,
       height = 8)
```

# Optimising Rstudio setup

# Rendering reports

You can also automate the running of Rmd files, to generate outputs

```{r}
rmarkdown::render("Analysis.Rmd", output_file = "outputs/Analysis.html")
```

# R environments




