---
title: "Class Notes on Reproducible Research"
output: html_document
date: "2024-07-24"
editor_options: 
  chunk_output_type: console
---


# What is ideal reproducible research?

- Transparent methods
- Accessible data and code
- Can be replicated
- Interoperable across operating systems and contexts

Historically, science was secretive but there has been a recent push to make science more open.

Many aspects of science can be made transparent, open, replicable. This course is about computation reproducibility.

Goal as an open scientist: Be able to reproduce the results at a later time, better if collaborator can do so, someone from research community.

There are processes to handle sensitive data (keep the science part open but the sensitive parts, private)

## Activity: Identify challenges/problems/barriers for reproducible analyses and collaboration

(see detailed notes in attached file)

- Software environment
- Workflow
- Coding conventions
- Versions
- Documentation
- Data management
- Big data

# Organisation

# Data curation and storage

```{r}

```



# R / Rmd / Qmd

# Reproducible coding

# Saving outputs

```{r}
library(tidyverse)
library(ggplot2)
library(maps)
library(ggthemes)

af <- read_csv("data/penguin_distributions/Aptenodytes forsteri.csv")

# Get world map data
world_map <- map_data("world")

af$countdatapresent<- case_when (af$individualCount>0 ~ "Count data present",
              .default = "No count data")

#mostly from chagpt:
# Create a ggplot object
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon( color = "white") + # Fill the regions and outline with white
  scale_fill_viridis_d() + # Use a viridis color scale for regions
  theme_minimal() + # Apply a minimal theme
  theme(
    panel.background = element_rect(fill = "lightblue"), # Light blue background
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    axis.title = element_blank(), # Remove axis titles
    axis.text = element_blank(), # Remove axis text
    axis.ticks = element_blank() # Remove the legend
  ) +
  coord_fixed(1.3)+
  geom_point(data=af, aes(x=decimalLongitude,y=decimalLatitude,group=NA,col=countdatapresent))# Ensure the aspect ratio is correct

ggsave("output/emperorpenguindist.pdf",width=5,height=8)
```

# Optimising Rstudio setup

# R environments


